Bioverse: Origins of Life

### 2023-11-06 src/tex/ms.tex
Restructure: one big Methods section.
No separate Survey Sims section.
highlight EECs
add short results, discussion, conclusion on spT dependency

\section{Hypothesis Tests}
\label{sec:hypotests}
\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}

@@

\subsection{Bayesian hypothesis testing with \bioverse}

@@

\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hz_nuv_evo.pdf}
        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass. White markers show age and host star mass of the transiting planets in the synthetic sample.
        Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
        \label{fig:hz_nuv_evo}
    \end{centering}
\end{figure*}

\section{Exoplanet survey simulations}
\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
% detecting methanogenic life: Sandora2020, Wang2018,
A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biosignature for our spectroscopic survey simulations.

\subsection{Habitable Worlds Observatory}
\subsubsection{target list}
\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts

\subsection{Nautilus}
\subsection{LIFE} % optional

@@

\subsection{Correlation of past UV flux and biosignature occurrence}

@@

\subsection{Contextual support for potential biosignature detections}
\todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}

@@

>>>>

\section{Methods}
\label{sec:hypotests}
\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
\subsection{Bayesian hypothesis testing with \bioverse}

@@

\subsection{Exoplanet survey simulations}
\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
% detecting methanogenic life: Sandora2020, Wang2018,
A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biosignature for our spectroscopic survey simulations.

\subsubsection{Habitable Worlds Observatory}
\subsubsection{target list}
\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts

\subsubsection{Nautilus}
\subsubsection{LIFE} % optional

@@

\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hz_nuv_evo.pdf}
        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass. The scatter plot shows age and host star mass of the transiting planets in the synthetic sample with \gls{EEC} highlighted in green.
        Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
        \label{fig:hz_nuv_evo}
    \end{centering}
\end{figure*}

@@

\subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}

In a magnitude- and volume-limited sample of a transit survey, the host star distribution will be skewed toward later spectral types and dominated by M~dwarfs (see Fig.~\ref{fig:inhabited_FGKM}).
Due to how the \gls{HZ} scales with spectral type, by far most transiting \glspl{EEC} occur around M~dwarfs.
These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Fig.~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
Because of that, all inhabited planets orbit M~dwarfs.

@@

\subsection{Contextual support for potential biosignature detections}
\todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
As we have shown in Section~\ref{sec:results_uv}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
This strong predicted correlation between stellar spectral type and the occurrence of life can be used to falsify \gls{ool} scenarios.
The constraint of this context is particularly strong if a candidate biosignature is detected on a planet orbiting an earlier type star, i.e., where it is unexpected in the context of this \gls{ool} scenario.

@@

Our main findings are:
\begin{enumerate}
    \item Abiogenesis in the cyanosulfidic scenario strongly prefers late host star spectral types in a sample of transiting planets. Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario, and any \gls{ool} theory requiring extended periods of high \gls{NUV} irradiation.
\end{enumerate}
